FraCaS: Temporal Analysis

### 2020-05-26 paper2020/paper.tex
minor

\section{Introduction}
Stergios
The semantics of tense and aspect has been a long standing issue in the study of formal semantics since the early days of Montague Grammar and a number of different ideas have been put forth on the table to deal with them throughout the years \citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The semantics of tense and aspect has been also central in the study of Natural Language Inference (NLI). The various datasets for NLI that have been proposed by the years contain examples that have some implicit or explicit reliance to inferences related to tense and aspect. One of the early datasets used to test logical approaches, the FraCaS test suite \citet{cooper:1996} contains a whole section of the dataset only for temporal and aspectual inference. This part of the FraCaS has been notoriously difficult to tackle. In fact, all the computational systems ran on the FraCaS so far skip this section, and usually also the anaphora and ellipsis sections.The exception with respect to the ellipsis and anaphora sections is the system presented by \citet{bernardy_type_2017,bernardy_wide_2019} where a whole run of the FraCaS excluding the temporal section has been attempted. In this paper, we take up the challenge of providing a computationally viable account of tense and aspect to deal with the notorious section 8 of the FraCaS test suite. This account is not meant to be a theoretically full account of tense and aspect, but an account that is driven by the need to cover the test suite in a way that is general enough to capture the test suite examples. The account is evaluated on the entailment properties of various temporal and aspectual examples, as given by the test suite. As such, we are not getting into the discussion of how tense and aspect might affect grammaticality or infelicitousness of various sentences. We take the sentences of the FraCaS to be syntactically correct and the entailment patterns they are annotated with valid.

@@

Unicity of action appears to be a non-logical principle. Indeed, it is
quite possible that ``Jones left'' several times. However, it seems
that this principle is never contradicted by the testsuite. As such,
even though unicity of action is only a pragmatic rule, it can be
taken as a valid one \emph{by default}: it is only when we have a
sufficiently constrained situation that one should reject it. Consider
the following discourse:
\begin{enumerate}[parsep=0pt,itemsep=1pt,label=(\arabic*)]
\item Smith left at 1pm.
\item Smith went to its appointment with the lawyer.
\item Smith left at 4pm.
\end{enumerate}
One would normally not say that there is contradiction. However if the
middle sentence were not present, a contradiction should be
flagged. We leave such discourse analysis as future work, and simply
apply unicity of action everywhere.

@@

\fracasex{314}
{\item	Smith arrived in Paris on the 5th of May, 1995.
\item	Today is the 15th of May, 1995.
\item	She is still in Paris.
\hyp 	Was Smith in Paris on the 7th of May, 1995?
}

@@

\fracasex{279}{
\item	Smith wrote a novel in 1991.
\hyp 	Did Smith write it in 1992?
}
Indeed, following our account, the above (contradictory) inference problem is to be
interperted as

@@

\fracasex{280}{
\item	Smith wrote a novel in 1991.
\item 	Did Smith write a novel in 1992?
\hyp 	Smith wrote a novel in 1992.
}
Here our interpretation is:

@@

% \bibliography{../PaperTools/bibtex/jp}

\end{document}

>>>>

\section{Introduction}
Stergios
The semantics of tense and aspect has been a long standing issue in the study of formal semantics since the early days of Montague Grammar and a number of different ideas have been put forth on the table to deal with them throughout the years \citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The semantics of tense and aspect has been also central in the study of Natural Language Inference (NLI). The various datasets for NLI that have been proposed by the years contain examples that have some implicit or explicit reliance to inferences related to tense and aspect. One of the early datasets used to test logical approaches, the FraCaS test suite \citet{cooper:1996} contains a whole section of the dataset only for temporal and aspectual inference (section 7). This part of the FraCaS has been notoriously difficult to tackle. In fact, all the computational systems ran on the FraCaS so far skip this section, and usually also the anaphora and ellipsis sections.The exception with respect to the ellipsis and anaphora sections is the system presented by \citet{bernardy_type_2017,bernardy_wide_2019} where a whole run of the FraCaS excluding the temporal section has been attempted. In this paper, we take up the challenge of providing a computationally viable account of tense and aspect to deal with the notorious section 8 of the FraCaS test suite. This account is not meant to be a theoretically full account of tense and aspect, but an account that is driven by the need to cover the test suite in a way that is general enough to capture the test suite examples. The account is evaluated on the entailment properties of various temporal and aspectual examples, as given by the test suite. As such, we are not getting into the discussion of how tense and aspect might affect grammaticality or infelicitousness of various sentences. We take the sentences of the FraCaS to be syntactically correct and the entailment patterns they are annotated with valid.

@@

Unicity of action appears to be a non-logical principle. Indeed, it is
quite possible that ``Jones left'' several times. However, it seems
that this principle is never contradicted by the testsuite. As such,
even though unicity of action is only a pragmatic rule, it can be
taken as a valid one \emph{by default}: it is only when we have a
sufficiently constrained situation that one should reject it. Consider
the following discourse:
\begin{enumerate}[parsep=0pt,itemsep=1pt,label=(\arabic*)]
\item Smith left at 1pm.
\item Smith went to his appointment with the lawyer.
\item Smith left at 4pm.
\end{enumerate}
One would normally not say that there is contradiction. However if the
middle sentence were not present, a contradiction should be
flagged. We leave such discourse analysis as future work, and simply
apply unicity of action everywhere.

@@

\fracasex{314}
{\item	Smith arrived in Paris on the 5th of May, 1995.
\item	Today is the 15th of May, 1995.
\item	She is still in Paris.
\hyp 	Smith was in Paris on the 7th of May, 1995.
}

@@

\fracasex{279}{
\item	Smith wrote a novel in 1991.
\hyp 	 Smith wrote it in 1992.
}
Indeed, following our account, the above (contradictory) inference problem is to be
interperted as

@@

\fracasex{280}{
\item	Smith wrote a novel in 1991.
\item 	Did Smith write a novel in 1992?
\hyp 	Smith wrote a novel in 1992.
}\todo{something wrong with P2 here, is there a premise there at all?}
Here our interpretation is:

@@

% \bibliography{../PaperTools/bibtex/jp}
% Things we cannot handle (I think): John was writing a book in 1991 -> John wrote a book in 1991 (I think we predict entailment here). This is because there is no distinction with progressive and non progressive aspect. We did not do it, I guess because it was needed, but just for future reference. There are similar examples playing around with LEXICAL aspect (aspect grammatically encoded) that we could look at in future work
% Accomplishments usually involve a culmination point in the action which is necessary for the event to be completed. This is not necessary for activities. John stopped writing a book not-> John wrote a book (but does imply that John was writing a book), but, John stopped running -> John run and John was runnning 
\end{document}
